---
title: "VUG in R"
author: "Michal Michalski"
date: "29/05/2020"
output: 
  html_document:
    theme: flatly
    highlight: tango
    code_download: true
    toc: true
    toc_float: 
      collapsed: true
      smooth_scroll: false
  
---
<style>
div.blue { background-color:#D3D3D3; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
Visualising Urban Geographies was a project by Edinburgh University that iprovided an amazing collection of geo-referenced maps, geographic boundaries and socio econoic maps - [Project Website](https://geo.nls.uk/urbhist/).
Here I use some of them to create a vizualizations in R.
</div>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r packages, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(sf)
library(ggplot2)
library(gganimate)
library(ggpomological)
library(magick)
library(dplyr)
library(readr)
library(ggmap)

```

## Chronological Map of Edinburgh
Chronological Map of Edinburgh Showing the Expansion of the City 1919 created by John George Bartholomew.

### Data

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
# load the shp
chrono_map = st_read("data/spatial/bart_chrono_1919.shp", quiet = TRUE)

#period colors
period_colors = c("#851613","#C83B42","#DE7B80","#F0AEB1",
                  "#135D89","#4D95BA","#96D1EA","#52442F","#BEB3A1","#F0E7D9")
#period labels
period_labels = c("pre 1450","1450 - 1515","1515 - 1622","1622 - 1750",
                  "1750 - 1800 ","1800 - 1825","1825 - 1850","1850 - 1875",
                  "1875 - 1900","since 1900")
```

### Animation

```{r message=FALSE, warning=FALSE, paged.print=FALSE, echo=TRUE}

#plot map for animation
map = ggplot() + 
  geom_sf(
    data = chrono_map,
    aes(fill= as.factor(End_Date)),
    alpha = 0.75, 
    colour = NA) +
  scale_fill_manual(name = "End of {current_frame}", values = period_colors, labels = period_labels) +
  theme_pomological() +
  labs(
    title = "Chronological Map of Edinburgh",
    subtitle = "Showing expansion of the City from earliest days to the present \nby J.G. Bartholomew - Cartographer to the King",
    caption = "https://geo.nls.uk/urbhist/"
  ) +
  transition_manual(End_Date, cumulative = T)

# plot static map
map = ggplot() + 
  geom_sf(
    data = chrono_map,
    aes(fill= as.factor(End_Date)),
    alpha = 0.75, 
    colour = NA) +
  scale_fill_manual(name = "period ", values = period_colors, labels = period_labels,
                    guide = guide_legend(
                      direction = 'horizontal',
                      title.position = 'top',
                      title.hjust = .5,
                      label.position = 'bottom',
                      label.hjust = .5,
                      keywidth = .5,
                      keyheight = .5,
                      nrow = 1
                    ))+
  labs(
    title = "Chronological Map of Edinburgh",
    subtitle = "Showing expansion of the City from earliest days to the present \nby J.G. Bartholomew - Cartographer to the King",
    caption = "Source: https://geo.nls.uk/urbhist/"
  ) +
  theme_pomological() +
  theme(
    legend.position = "bottom"
  ) +
  transition_manual(End_Date, cumulative = T)

#create animation
animate(map, width = 800, height = 600)

#save gif
anim_save("map_anim.gif", map, width = 800,height = 600)

```

## Occupations
Selection of Edinburgh profession from 1911.
```{r}
occupations_edi = read_csv("data/processed/occupations.csv")

base_plot = qmplot(
  data = occupations_edi,
  x = long,
  y = lat,
  zoom = 14,
  geom = "blank",
  maptype = "toner-lines",
  source = 'stamen'
)

base_plot +
  geom_point(mapping = aes(x = long, y = lat), size = .5) +
  labs(
    title = "Edinburgh Occupations in 1911",
    caption = "Source: vug | @topographos2"
  ) +
  theme_pomological(
  ) +
  theme(
    axis.title.x=element_blank(),
    axis.title.y=element_blank()
  ) +
  facet_wrap(~ occupation)
```



```{r}
sessionInfo()
```


&nbsp;
<hr />
<p style="text-align: center;">A work by <a href="https://github.com/topographos/">Michal Michalski</a></p>
<p style="text-align: center;"><span style="color: #808080;"><em>michal.m.michalski@durham.ac.uk</em></span></p>
